An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-08-02 EconGrowth2.Rnw
Added subsection on CES parameter estimation.

>>>>

%++++++++++++++++++++++++++++++
\subsection{CES Models}
\label{sec:CES_Parameter_Estimation}
%++++++++++++++++++++++++++++++

The CES model without and with energy is given by 
Equations~\ref{eq:CES}--\ref{eq:CESekl}. 
The \texttt{R} \citep{R} package \texttt{micEconCES} 
\citep{Henningsen:2011td} was used to estimate parameters 
$\gamma$, $\lambda$, $\delta_1$, $\rho_1$, $\delta$, and $\rho$. 
**** Adjust these parameter names if we decide to 
change them. **** The \texttt{cesEst} function of \texttt{micEconCES} provides 
several algorithm options for parameter estimation.
The default algorithm (Levenberg-Marquardt) does not respect
parameter constraints (see Section~\ref{sec:CES}) and, 
in our testing, nearly always violated them, often returning
negative values for elasticity of substitution ($\sigma$) parameters.

Two algorithms available in \texttt{cesEst} 
(\texttt{PORT} and \texttt{L-BFGS-B}) respect coefficient constraints. 
Our CES parameter estimation algorithm tries both
with default starting values and selects the
result with lowest $sse$, regardless
of convergence status. If one algorithm
fails during the fitting process 
(due to division by zero or another numerical error), 
we use coefficients from the algorithm that didn't.
In event that both algorithms fail, 
we skip that resampled data point.

\citet{Henningsen:2011td}, 
in their detailed analysis of \citet{Kemfert:1998vd}, found that

\begin{quote}
\ldots the Levenberg-Marquardt and the PORT algorithms 
are---at least in this study---most likely to find the 
coefficients that give the best fit to the model, 
where the PORT algorithm can be used to restrict the estimates 
to the economically meaningful region.
\end{quote}

In our testing, we found that to be mostly true. 
\texttt{PORT} nearly always provided lower
$sse$ than \texttt{L-BFGS-B}, despite the fact that 
\texttt{L-BFGS-B} often reports convergence
and \texttt{PORT} does not for the same economic data.

**** Include CES graphs here. Need to cover
all options: without energy and the three nesting
options. CIs? TBD which to include. ****

Figure~\ref{fig:CES_GDP_Lattice_Graph_kle} compares CES model predictions 
to historical data for the $(kl)e$ nesting.

<<CES_GDP_Lattice_Graph_kle, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="CES fitted models with ($kl$)$e$ nesting.">>=
createCESLatticeGraph()
@

**** Include other nestings here? ****
